---
title: "KZ_HW04"
output: 
    github_document:
        toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Welcome to my 4th homework assignment. 
In this assignment I will complete two tasks. 
For my first task I will change the shape of my data and for the second I will combine two data sources into one data object.

# Task #1
## Activity 2: 

Make a tibble with one row per year and columns for life expectancy for two or more countries.
Use knitr::kable() to make this table look pretty in your rendered homework.
Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.

For this assignment I will use the **gapminder** dataset. I will begin by loading the dataset as well as packages I will be using (such as the knitr package to use the kable function)

```{r}
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(kableExtra))
```

For this task I plan to use the *gather* and *spread* functions. Before I begin let's talk about what these functions are.

**Gather** -> takes multiple columns and gather them into key-value pairs (makes wide data longer OR making untidy data tidy)

**Spread** -> takes a key-valye pair and spreads them into multiple columns (makes long data wider OR making tidy data untidy to please the eye)

Tyler Rinker has a great guide for spread and gather functions at this [link](https://github.com/trinker/tidyr_in_a_nutshell/blob/master/README.md)

Remember: From previous exercises, we can also use the **select** function to choose certain columns and **filter** function to select certain rows.

```{r}
#our first step is to create a tibble with one row per year and columns for life expectancy for two or more countries.

#Gapminder is already a tibble but we are going to create changes to it so we will rename it as a new tibble

#this new tibble will be called *new_gap*

new_gap <- gapminder %>%
  filter(country %in% c('Canada', 'Australia', 'New Zealand')) %>%
          select(country, year, lifeExp) %>% #Now I have created a new tibble for New Zealand and Canada's life expectancy. 
# My Next move is to create a column for each country. For this I will use the *spread* function
spread(country, lifeExp) 

new_gap %>%
   knitr::kable(align = "c", caption = "Canada, Australia, and New Zealand's Life Expectancy Per Year")


#This gives us a table giving the year, each country, and the life expectancy for each country. Next I will make a scatterplot of life expectancy for Canada against New Zealand

new_gap %>% 
  ggplot(aes(Canada, Australia)) +
  geom_jitter() +
  geom_smooth(method = "loess") +
  geom_point() +
  ggtitle("Canada vs Australia life expectancy") 

```

# Task #2
## Activity 1: 
Create a second data frame, complementary to Gapminder. Join this with (part of) Gapminder using a dplyr join function and make some observations about the process and result. Explore the different types of joins. Examples of a second data frame you could build:
One row per country, a country variable and one or more variables with extra info, such as language spoken, NATO membership, national animal, or capitol city.
One row per continent, a continent variable and one or more variables with extra info, such as northern versus southern hemisphere.

I will accomplish this task by using gapminder (a pre-existing dataset) and a dataset that I created, by importing it.


```{r}
#Now I will upload and view my data
country_animals <- read.csv("~/Documents/country_nat_animals.csv")

#Next, I will put my data in a table so you can see it 
country_animals %>%
  knitr::kable(align = "c")
```

Okie dokie. My next mission is to join this with the gapminder dataset. I will combine the gapminder dataset and the national animals column from the country_animals dataset. 

```{r}
gapminder_animals <- country_animals %>% 
  select(country, National.Animal)

#What I will do next is test out the different types of join functions

#Let's start with an inner join:
gapminder_animals %>%
  inner_join(gapminder, national.animal, by = "country") %>%
  knitr::kable(align = "c")
```
### inner_join(x, y): Return all rows from x where there are matching values in y, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned. This is a mutating join.

```{r}
#Next, let's try a semi-join:
gapminder_animals %>%
  semi_join(gapminder, national.animal, by = "country") %>%
  knitr::kable(align = "c")
```

### semi_join(x, y): Return all rows from x where there are matching values in y, keeping just columns from x. A semi join differs from an inner join because an inner join will return one row of x for each matching row of y, where a semi join will never duplicate rows of x. This is a filtering join.

```{r}
#Next, let's try a left join:
gapminder_animals %>%
  left_join(gapminder, national.animal, by = "country") %>%
  knitr::kable(align = "c")
```

### left_join(x, y): Return all rows from x, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned. This is a mutating join.

```{r}
#Next, let's try an anti join:
gapminder_animals %>%
  anti_join(gapminder, national.animal, by = "country") %>%
  knitr::kable(align = "c")
```

### anti_join(x, y): Return all rows from x where there are not matching values in y, keeping just columns from x. This is a filtering join.

```{r}
#Next, let's try a full join:
gapminder_animals %>%
  full_join(gapminder, national.animal, by = "country") %>%
  knitr::kable(align = "c")
```

### full_join(x, y): Return all rows and all columns from both x and y. Where there are not matching values, returns NA for the one missing. This is a mutating join.

After running through the scenarios, it is clear that some types of the join function were more functional for my dataset than others. The full_join was not beneficial because it spat out all entries, even when there were NA values for countries that did not have a naitonal animal input. Inner join did not look that great either as it resulted in multiple entries for each country. There are clearly different types of join functions to suit your needs. 

> I want to mention that I wanted to make my tables scroll tables, but that would require me to change my output to HTML. I wanted to leave it as a github document. Cheers!