---
title: "KZ_HW04"
output: 
    github_document:
        toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Welcome to my 4th homework assignment. 
In this assignment I will complete two tasks. 
For my first task I will change the shape of my data and for the second I will combine two data sources into one data object.

# Task #1
## Activity 2: 

Make a tibble with one row per year and columns for life expectancy for two or more countries.
Use knitr::kable() to make this table look pretty in your rendered homework.
Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.

For this assignment I will use the **gapminder** dataset. I will begin by loading the dataset as well as packages I will be using (such as the knitr package to use the kable function)

```{r}
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(kableExtra))
```

For this task I plan to use the *gather* and *spread* functions. Before I begin let's talk about what these functions are.

**Gather** -> takes multiple columns and gather them into key-value pairs (makes wide data longer OR making untidy data tidy)

**Spread** -> takes a key-valye pair and spreads them into multiple columns (makes long data wider OR making tidy data untidy to please the eye)

Tyler Rinker has a great guide for spread and gather functions at this [link](https://github.com/trinker/tidyr_in_a_nutshell/blob/master/README.md)

Remember: From previous exercises, we can also use the **select** function to choose certain columns and **filter** function to select certain rows.

```{r}
#our first step is to create a tibble with one row per year and columns for life expectancy for two or more countries.

#Gapminder is already a tibble but we are going to create changes to it so we will rename it as a new tibble

#this new tibble will be called *new_gap*

new_gap <- gapminder %>%
  filter(country %in% c('Canada', 'Australia', 'New Zealand')) %>%
          select(country, year, lifeExp) %>% #Now I have created a new tibble for New Zealand and Canada's life expectancy. 
# My Next move is to create a column for each country. For this I will use the *spread* function
spread(country, lifeExp) 

new_gap %>%
   knitr::kable(align = "c", caption = "Canada, Australia, and New Zealand's Life Expectancy Per Year")


#This gives us a table giving the year, each country, and the life expectancy for each country. Next I will make a scatterplot of life expectancy for Canada against New Zealand

new_gap %>% 
  ggplot(aes(Canada, Australia)) +
  geom_jitter() +
  geom_smooth(method = "loess") +
  geom_point() +
  ggtitle("Canada vs Australia life expectancy") 

```

Please see the 2nd part of my assignment in **part 2** linked [here]() and in my README file. The file was too big to be in one so I split it up into two.

